Review Paper Draft

### 2020-06-25 main.tex
Figure updates - in progress

\pb{I don't like calling these two methods ``equivalent''.  It immediately begs the question: if they're equivalent, why is one of the main aims of this paper to compare them?  I propose renaming this section ``Relationship between PC-Lasso and LMM-Lasso'' and then talking about both their similarities AND their differences.}

Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \ref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects. \cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \ref{eqn:k}. \pb{use \textbackslash eqref, not \textbackslash ref, for equations}

@@

\pb{should this paragraph move to next section?}
Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
\begin{equation}
    \label{eqn:sim}
    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
\end{equation}
where $\by$ represents the phenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
\begin{equation}
    \label{eqn:parition_y}
    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
\end{equation}
Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise. \pb{So, for example, $\eta=0.5$ and $xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).

@@

\anna{
\begin{itemize}
    \item Need details about the real data once it's run
    \item Should I mention subpopulation sizes (a few large ones) or gamma structure (half-and-half-heterogeneous) here? No - do a brief summary where noted above
\end{itemize}
}

\begin{figure}[H]
    \centering
    \includegraphics[scale = 1.1]{figures/beta_mse.png}
    \caption{MSE for different data types in the presence and absence of environmental confounding.}
    \label{fig:mse}
\end{figure}

@@

\anna{
\begin{itemize}
    \item I ran this but think it just makes things more confusing - go back to only independent subpopulations data - for both (?) 
\end{itemize}
}

\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/fig2a.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy.}
    \label{fig:big_vs_small}
\end{figure}

@@

\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/fig2b.png}
    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
    \label{fig:big_small_gamma}
\end{figure}

@@

\anna{eta part needs to be converted to mse}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/eta_beta_hat.png}
\end{figure}

\newpage
\anna{The error bars are indicating differences across different values of xi and simulated data-types (indepent and 1d linear admixture - this may be confusing/something to change or describe explicitly.
}

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.8]{figures/time_ratio.png}
    \caption{Caption goes here}
    \label{fig:time}
\end{figure}

In addition, we have found that LMM-Lasso-ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the RRM. Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners. However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in \anna{figure xxx}, this is also a setting in which the computational time required may be prohibitive for practical applications. Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes. The horizontal dashed line at 1 indicates both methods require the same computing time. Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes approximately 71 to 9 times as long as LMM-Lasso-Rakitsch. Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.

>>>>

\pb{I don't like calling these two methods ``equivalent''.  It immediately begs the question: if they're equivalent, why is one of the main aims of this paper to compare them?  I propose renaming this section ``Relationship between PC-Lasso and LMM-Lasso'' and then talking about both their similarities AND their differences.} \anna{sounds good}

Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \ref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects. \cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}. \pb{use \textbackslash eqref, not \textbackslash ref, for equations} \anna{ok, but why?}

@@

\pb{should this paragraph move to next section?}
\anna{yes} Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
\begin{equation}
    \label{eqn:sim}
    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
\end{equation}
where $\by$ represents the phenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
\begin{equation}
    \label{eqn:parition_y}
    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
\end{equation}
Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise. \pb{So, for example, $\eta=0.5$ and $xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).

@@

\anna{
\begin{itemize}
    \item Need details about the real data once it's run
    \item Need to fix facet labels
\end{itemize}
}

\begin{figure}[H]
    \centering
    \includegraphics[scale = 1.1]{figures/beta_mse.png}
    \caption{MSE for different data types in the presence and absence of environmental confounding. \anna{$\gamma$ structure is half-and-half-decreasing-heterogeneous, subpop size is large}}
    \label{fig:mse}
\end{figure}

@@

\anna{admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g. 4 environmental exposures. the small subpopulations scenario more closely mimics gradations of environmental effects - what is usually simulated with a random draw from the MVN distn... we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}

\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy:  \anna{$\bgamma$ structure is nonlinear (aka half and half) and heterogeneous, $\eta= \xi = 0.5$}}
    \label{fig:big_vs_small}
\end{figure}

@@

\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/fig2b.png}
    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures: \anna{structureX is 1d linear admixture, $\eta= \xi = 0.5$}}
    \label{fig:big_small_gamma}
\end{figure}

@@

\anna{Fix factor label greek letters - R update broke some things}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/eta_beta_hat.png}
     \caption{Caption goes here \anna{$\xi = 0.5$, structureX = 1d linear, structureGamma is half and half heterogeneous}}
    \label{fig:eta_beta_mse}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.8]{figures/time_ratio.png}
    \caption{Caption goes here \anna{$\eta = \xi = 0.5$, structureX = 1d linear, structureGamma is half and half heterogeneous}}
    \label{fig:time}
\end{figure}

In addition, we have found that LMM-Lasso-ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the RRM. Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners. However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications. Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes. The horizontal dashed line at 1 indicates both methods require the same computing time. Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes approximately 71 to 9 times as long as LMM-Lasso-Rakitsch. Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
